\textit{Extended data for}: Towards human Super EEG?

### 2018-10-11 paper/main.tex
edits to paper and figures

\newcommand{\perexpcorrmaps}{S1}
\newcommand{\perexptaskrecon}{S2}
\newcommand{\perexptaskreconseparated}{S3}
\newcommand{\suppstats}{S4}

@@

\section*{Approach}
The SuperEEG approach to inferring high temporal resolution full-brain
activity patterns is outlined and summarized in
Figure~\ref{fig:methods}. We describe (in this section) and evaluate
(in \textit{Results}) our approach using a two large previously
collected dataset comprising multi-session intracranial recordings.
Dataset 1 comprises multi-session recordings taken from 6876
electrodes implanted in the brains of 88 epilepsy
patients~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12}.  Each recording session lasted from 0.3--14.2 hours
(Fig.~\suppstats A), and includes data recorded roughly from when the
patients woke up each morning, to before they went to sleep at the end
of each day.  In addition to typical bed-ridden hospital patient
activities (e.g., lying in bed, reading, watching television, using
personal electronic devices, listening to music, visiting with family
and friends, etc.), the patients also performed a variety of
experimental cognitive tasks throughout their day (primarily
list-learning memory tasks).  For the purposes of the Dataset 1
analyses presented here, we aggregated all data across each recording
session, ignoring the particular activities or tasks the patients were
performing at any given moment.  We used Dataset 1 to develop our main
SuperEEG approach, and to examine the extent to which SuperEEG might
be able to generate task-general predictions.  Dataset 2 comprised
multi-session recordings from 3159 electrodes implanted in the brains
of 24 epilepsy patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17,
  KuceEtal17, LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18,
  KuceEtal18}.  Each recording session lasted from .4--6.6 hours
(Fig.~\suppstats B).  Whereas Dataset 1 included recordings taken during a
wide variety of behaviors, Dataset 2 included recordings taken as each
patient performed each of two memory tasks: a random word list free
recall task and a categorized word list free recall task.  We used
Dataset 2 to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.  Figure~\suppstats~provides
additional details on both datasets.

@@

To illustrate our approach, we first examine an individual held-out
raw LFP trace and its associated SuperEEG-derived reconstruction.
Figure~\ref{fig:recon}A displays the observed LFP from the red
electrode in Figure~\ref{fig:methods}A (blue), and its associated
reconstruction (red), during a 5~s time window during one of the
example patient's six recording sessions.  The two traces match
closely ($r = 0.869, p < 10^{-10}$).  Figure~\ref{fig:recon}B displays a
two-dimensional histogram of the actual versus reconstructed voltages
for the entire 14.2 total hours of recordings from the example
electrode (correlation: $r = 0.909, p < 10^{-10}$).  This example confirms
that the SuperEEG algorithm recovers the recordings from this single
electrode well.  Next, we used this general approach to quantify the
algorithm's performance across the full dataset.

@@

For each held-out electrode, from each held-out patient in turn, we
computed the average correlation (across recording sessions) between
the SuperEEG-reconstructed voltage traces and the observed voltage
traces from that electrode.  For this analysis we set $\bar{R}$ to be
the union of all electrode locations across all patients.  This
yielded a single correlation coefficient for each electrode location
in $\bar{R}$, reflecting how well the SuperEEG algorithm was able to
recover the recording at that location by incorporating data across
patients (black histogram in Fig.~\ref{fig:corrmap}A, map in
Fig.~\ref{fig:corrmap}C).  The observed distribution of correlations
was centered well above zero (mean: 0.52; $t$-test comparing mean of
distribution of $z$-transformed per-electrode correlation coefficients
to 0: $t(66) = 25.08, p < 10^{-30}$), indicating that the SuperEEG algorithm
recovers held-out activity patterns substantially better than random
guessing.

As a stricter benchmark, we compared the quality of these
across-participant reconstructions (i.e., computed using a correlation
model learned from other patients' data) to reconstructions generated
using a correlation model trained using the in-patient's data.  In
other words, for this within-patient benchmark analysis we estimated
$\hat{C}_{s}$ (Eqn.~\ref{eqn:subj_corrmat}) for each patient in turn,
using recordings from all of that patient's electrodes except at the
location we were reconstructing.  These within-patient reconstructions
serve as an estimate of how well data from all of the other electrodes
from that single patient may be used to estimate held-out data from
the same patient.  This allows us to ask how much information about
the activity at a given electrode might be inferred through (a) volume
conductance or other sources of ``leakage'' from activity patterns
measured from the patient's other electrodes and (b) across-electrode
correlations learned from that single patient.  As shown in
Figure~\ref{fig:corrmap}A (gray histogram), the distribution of
within-patient correlations was centered well above zero (mean: 0.32;
$t$-test comparing mean of distribution of $z$-transformed
per-electrode correlation coefficients to 0: $t(66) = 15.16, p < 10^{-20}$).
However, the across-patient correlations were substantially higher
($t$-test comparing average $z$-transformed within versus across
patient electrode correlations: $t(66) = 9.62, p < 10^{-10}$).  This is an
especially conservative test, given that the across-patient SuperEEG
reconstructions exclude (from the correlation matrix estimates) all
data from the patient whose data is being reconstructed.  We repeated
each of these analyses on a second independent dataset and found
similar results (Fig.~\ref{fig:corrmap}B, D; within versus across
reconstruction accuracy: $t(23) = 6.93, p < 10^{-5}$). We also
replicated this result separately for each of the two experiments from
Dataset 2 (Fig.~\perexpcorrmaps).  This overall finding, that
reconstructions of held-out data using correlation models learned from
\textit{other} patient's data yield higher reconstruction accuracy
than correlation models learned from the patient whose data is being
reconstructed, has two important implications.  First, it implies that
distant electrodes provide additional predictive power to the data
reconstructions beyond the information contained solely in nearby
electrodes.  (This follows from the fact that each patient's
electrodes are implanted in a unique set of locations, so for any
given electrode the closest electrodes in the full dataset are likely
to come from the same patient.)  Second, it implies that the spatial
correlations learned using the SuperEEG algorithm are, to some extent,
similar across people.

@@

To gain insights into whether particular electrode locations might be
especially informative, we first computed the average reconstruction
accuracy across all of each patient's electrodes (using the
across-patients cross validation test; black histograms in
Fig.~\ref{fig:corrmap}A and B).  We labeled each patient's electrodes
in each dataset with the average reconstruction accuracy for that
patient.  In other words, we assigned every electrode from each given
patient the same value, reflecting how well the activity patterns at
those electrodes were reconstructed on average.  Next, for each voxel
in the 4~mm$^3$ MNI brain, we computed the average value across any
electrode (from any patient) that came within 20 MNI units of that
voxel's center.  Effectively, we computed an \textit{information
  score} for each voxel, reflecting the average reconstruction
accuracy across any patients with electrodes near each voxel-- where
the averages were weighted to reflect patients who had more electrodes
implanted near that location. This yielded a single map for each
dataset, highlighting regions that are potentially promising
implantation targets in terms of providing full-brain activity
information via SuperEEG (Fig.~\ref{fig:informap}A, B).  Despite task
and patient differences across the two datasets, we nonetheless found
that the maps of the most promising implantation targets derived from
both datasets were similar (voxelwise correlation between information
scores across the two datasets: $r = 0.23, p < 10^{-10}$).  While the
correspondence between the two maps was not perfect, our finding that
there were some commonalities between the two maps lends support to
the notion that different brain areas are differently informative
about full-brain activity patterns.  Further, regions like the right post
central gyrus, right supramarginal gyrus,
and the right nucleus accumbens that exhibited strong information scores in both datasets may
be especially promising implantation targets.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/informap}
  \caption{\textbf{Most informative electrode locations.}
    \textbf{A. Dataset 1 information score by voxel.} The voxel colors
    reflect the weighted average reconstruction accuracy across all
    electrodes from any patients with at least one electrode within 20
    MNI units of the given voxel.  \textbf{B. Dataset 2 information
      score by voxel.}  This panel is in the same format as Panel A.
    In Panels A and B the contours indicate the intersections between
    the top 10\% most informative voxels in each map.
    \textbf{C. Correspondence in information scores by voxel across
      Datasets 1 and 2.}  Same format as Figure~\ref{fig:density}C.}
  \label{fig:informap}
\end{figure}

@@

\section*{Code availability}
We have published an open-source toolbox implementing the SuperEEG
algorithm.  It may be downloaded \href{https://supereeg.readthedocs.io/en/latest/}{\underline{here}}.

>>>>

\newcommand{\perexpcorrmaps}{S1}
\newcommand{\perexptaskrecon}{S2}
\newcommand{\perexptaskreconseparated}{S3}
\newcommand{\suppstats}{S4}
\newcommand{\intersectmap}{S5}

@@

\section*{Approach}
The SuperEEG approach to inferring high temporal resolution full-brain
activity patterns is outlined and summarized in
Figure~\ref{fig:methods}. We describe (in this section) and evaluate
(in \textit{Results}) our approach using a two large previously
collected dataset comprising multi-session intracranial recordings.
Dataset 1 comprises multi-session recordings taken from 6876
electrodes implanted in the brains of 88 epilepsy
patients~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12}.  Each recording session lasted from {0.2}--{3.0}
(total recording time: {0.3}--14.2 hours; Fig.~\suppstats A), and includes data from a list-learning memory experiment
for up to one hour varying by patient with additional buffer time in
which the patients performed typical bed-ridden hospital
activities.  For the purposes of the Dataset 1
analyses presented here, we aggregated all data across each recording
session, ignoring the particular activities or tasks the patients were
performing at any given moment.  We used Dataset 1 to develop our main
SuperEEG approach, and to examine the extent to which SuperEEG might
be able to generate task-general predictions.  Dataset 2 comprised
multi-session recordings from 4436 electrodes implanted in the brains
of 40 epilepsy patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17,
  KuceEtal17, LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18,
  KuceEtal18}.  Each recording session lasted from {0.4}--2.2 hours
(total recording time: {0.4}--6.6 hours; Fig.~\suppstats B).  Whereas Dataset 1 included recordings taken during a
wide variety of behaviors, Dataset 2 included recordings taken as each
patient performed each of two memory tasks: a random word list free
recall task and a categorized word list free recall task.  We used
Dataset 2 to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.  Figure~\suppstats~provides
additional details on both datasets.

@@

To illustrate our approach, we first examine an individual held-out
raw LFP trace and its associated SuperEEG-derived reconstruction.
Figure~\ref{fig:recon}A displays the observed LFP from the red
electrode in Figure~\ref{fig:methods}A (blue), and its associated
reconstruction (red), during a 5~s time window during one of the
example patient's six recording sessions.  The two traces match
closely ($r = 0.86, p < 10^{-10}$).  Figure~\ref{fig:recon}B displays a
two-dimensional histogram of the actual versus reconstructed voltages
for the entire 14.2 total hours of recordings from the example
electrode (correlation: $r = 0.91, p < 10^{-10}$).  This example confirms
that the SuperEEG algorithm recovers the recordings from this single
electrode well.  Next, we used this general approach to quantify the
algorithm's performance across the full dataset.

@@

For each held-out electrode, from each held-out patient in turn, we
computed the average correlation (across recording sessions) between
the SuperEEG-reconstructed voltage traces and the observed voltage
traces from that electrode.  For this analysis we set $\bar{R}$ to be
the union of all electrode locations across all patients.  This
yielded a single correlation coefficient for each electrode location
in $\bar{R}$, reflecting how well the SuperEEG algorithm was able to
recover the recording at that location by incorporating data across
patients (black histogram in Fig.~\ref{fig:corrmap}A, map in
Fig.~\ref{fig:corrmap}C).  The observed distribution of correlations
was centered well above zero (mean: 0.52; $t$-test comparing mean of
distribution of $z$-transformed average patient correlation coefficients
to 0: $t(66) = 25.08, p < 10^{-10}$), indicating that the SuperEEG algorithm
recovers held-out activity patterns substantially better than random
guessing.

As a stricter benchmark, we compared the quality of these
across-participant reconstructions (i.e., computed using a correlation
model learned from other patients' data) to reconstructions generated
using a correlation model trained using the in-patient's data.  In
other words, for this within-patient benchmark analysis we estimated
$\hat{C}_{s}$ (Eqn.~\ref{eqn:subj_corrmat}) for each patient in turn,
using recordings from all of that patient's electrodes except at the
location we were reconstructing.  These within-patient reconstructions
serve as an estimate of how well data from all of the other electrodes
from that single patient may be used to estimate held-out data from
the same patient.  This allows us to ask how much information about
the activity at a given electrode might be inferred through (a) volume
conductance or other sources of ``leakage'' from activity patterns
measured from the patient's other electrodes and (b) across-electrode
correlations learned from that single patient.  As shown in
Figure~\ref{fig:corrmap}A (gray histogram), the distribution of
within-patient correlations was centered well above zero (mean: 0.32;
$t$-test comparing mean of distribution of $z$-transformed
average patient correlation coefficients to 0: $t(66) = 15.16, p < 10^{-10}$).
However, the across-patient correlations were substantially higher
($t$-test comparing average $z$-transformed within versus across
patient electrode correlations: $t(66) = 9.62, p < 10^{-10}$).  This is an
especially conservative test, given that the across-patient SuperEEG
reconstructions exclude (from the correlation matrix estimates) all
data from the patient whose data is being reconstructed.  We repeated
each of these analyses on a second independent dataset and found
similar results (Fig.~\ref{fig:corrmap}B, D; within versus across
reconstruction accuracy: $t(23) = 6.93, p < 10^{-5}$). We also
replicated this result separately for each of the two experiments from
Dataset 2 (Fig.~\perexpcorrmaps).  This overall finding, that
reconstructions of held-out data using correlation models learned from
\textit{other} patient's data yield higher reconstruction accuracy
than correlation models learned from the patient whose data is being
reconstructed, has two important implications.  First, it implies that
distant electrodes provide additional predictive power to the data
reconstructions beyond the information contained solely in nearby
electrodes.  (This follows from the fact that each patient's
electrodes are implanted in a unique set of locations, so for any
given electrode the closest electrodes in the full dataset are likely
to come from the same patient.)  Second, it implies that the spatial
correlations learned using the SuperEEG algorithm are, to some extent,
similar across people.

@@

To gain insights into whether particular electrode locations might be
especially informative, we first computed the average reconstruction
accuracy across all of each patient's electrodes (using the
across-patients cross validation test; black histograms in
Fig.~\ref{fig:corrmap}A and B).  We labeled each patient's electrodes
in each dataset with the average reconstruction accuracy for that
patient.  In other words, we assigned every electrode from each given
patient the same value, reflecting how well the activity patterns at
those electrodes were reconstructed on average.  Next, for each voxel
in the 4~mm$^3$ MNI brain, we computed the average value across any
electrode (from any patient) that came within 20 MNI units of that
voxel's center.  Effectively, we computed an \textit{information
  score} for each voxel, reflecting the average reconstruction
accuracy across any patients with electrodes near each voxel-- where
the averages were weighted to reflect patients who had more electrodes
implanted near that location. This yielded a single map for each
dataset, highlighting regions that are potentially promising
implantation targets in terms of providing full-brain activity
information via SuperEEG (Fig.~\ref{fig:informap}A, B).  Despite task
and patient differences across the two datasets, we nonetheless found
that the maps of the most promising implantation targets derived from
both datasets were similar (voxelwise correlation between information
scores across the two datasets: $r = 0.03, p < 10^{-10}$).  While the
correspondence between the two maps was not perfect, our finding that
there were some commonalities between the two maps lends support to
the notion that different brain areas are differently informative
about full-brain activity patterns.  Further, we found the
intersection of the top 10\% of the most informative voxels between
the two datasets (Fig.~\intersectmap) and found gray matter regions like the right postcentral gyrus, right supramarginal gyrus,
and the right nucleus accumbens that exhibited strong information scores in both datasets may
be especially promising implantation targets.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/informap}
  \caption{\textbf{Most informative electrode locations.}
    \textbf{A. Dataset 1 information score by voxel.} The voxel colors
    reflect the weighted average reconstruction accuracy across all
    electrodes from any patients with at least one electrode within 20
    MNI units of the given voxel.  \textbf{B. Dataset 2 information
      score by voxel.}  This panel is in the same format as Panel A.
    In both panels the contours indicate the intersections between
    the top 10\% most informative voxels in each map.
    \textbf{C. Correspondence in information scores by voxel across
      Datasets 1 and 2.}  Same format as Figure~\ref{fig:density}C.}
  \label{fig:informap}
\end{figure}

@@

\section*{Code availability}
We have published an open-source toolbox implementing the SuperEEG
algorithm.  It may be downloaded
\href{https://supereeg.readthedocs.io/en/latest/}{\underline{here}}. Additionally,
we have provided notebooks for all analyses and figures reported
here.

### 2018-10-11 paper/supporting_information.tex
add compiled supporting info pdf
edits to paper and figures

\title{\textit{Supporting Information for}: Towards human Super EEG?}
\author{Lucy L. W. Owen, Andrew C. Heusser, and Jeremy R. Manning\\Department of Psychological and Brain Sciences\\Dartmouth College, Hanover, NH 03755, USA\\Corresponding author: jeremy.r.manning@dartmouth.edu}

@@

\begin{figure}[p!]
\centering
\includegraphics[width=0.75\textwidth]{figs/supplemental_1}
\caption{\small \textbf{Reconstruction quality for datasets 1 and 2.} \textbf{A. Distributions
      of correlation coefficients.}  Although patients in dataset 1 all participated in the same experiment, we did not limit our analyses to task specific data.  Instead we incorporated data from their entire recording session (sometimes up to 14 hours of data) and classify this as training our model on across experiment information. The split violin plot shows the distributions of reconstruction accuracy using models trained on data from all other patients (Across Subject, in black) and all other electrodes from the same patient (Within Subject, in gray). \textbf{B.  Distributions of correlation coefficients.} Since all patients participated in two experiments in dataset 2, we limit our analyses to task specific data.  The violin plots compare distributions of reconstruction accuracy using data from the same task (Within Experiment), the other task (Across Experiment), and from both tasks (All Experiment), using models trained on data from all other patients (Across Subject, in black) and all other electrodes from the same patient (Within Subject, in gray).}
\label{fig:supplemental_1}
\end{figure}

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{figs/supplemental_2}
\caption{\small \textbf{Reconstruction quality for experiments 1 and 2.} \textbf{A. Distributions
      of correlation coefficients.} For experiment 1, the split violin plots compare distributions of reconstruction accuracy using data from just experiment 1 (Within Experiment), just experiment 2 (Across Experiment), and from both experiments (All Experiment), using models trained on data from all other patients (Across Subject, in black) and all other electrodes from the same patient (Within Subject, in gray). \textbf{B.  Distributions
      of correlation coefficients.} For experiment 2, the split violin plots compare distributions of reconstruction accuracy using data from just experiment 2 (Within Experiment), just experiment 1 (Across Experiment), and from both experiments (All Experiment), using models trained on data from all other patients (Across Subject, in black) and all other electrodes from the same patient (Within Subject, in gray). }
\label{fig:supplemental_2}
\end{figure}

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{figs/supplemental_3}
\caption{\small \textbf{Reconstruction quality for each experiment.} \textbf{A. \& C.  Distributions
      of correlation coefficients.}  Across all electrodes from all
    patients in the labeled experiment from dataset 2, the panel displays the distribution of correlations between the observed and reconstructed LFP data using models trained on data from all other patients (Across, in black) and all other electrodes from the same patient (Within, in gray). 
    \textbf{B. \& D. Correlation maps.}  The glass brain maps display the
    average correlation between the observed LFP data and the across-subjects model reconstructed data by location, for each labeled experiment.}
\label{fig:supplemental_3}
\end{figure}

>>>>

\title{\textit{Extended data for}: Towards human Super EEG?}
\author{Lucy L. W. Owen, Andrew C. Heusser, and Jeremy R. Manning\\Department of Psychological and Brain Sciences\\Dartmouth College, Hanover, NH 03755, USA\\Corresponding author: jeremy.r.manning@dartmouth.edu}

@@

\begin{figure}[phtb]
\centering
\includegraphics[width=0.75\textwidth]{figs/supplemental_1}
\caption{\small \textbf{Reconstruction quality for datasets 1 and 2.} \textbf{A. Distributions
      of correlation coefficients.}  Although patients in dataset 1 all participated in the same experiment, we did not limit our analyses to task specific data.  Instead we incorporated data from their entire recording session (sometimes up to 14 hours of data) and classify this as training our model on across experiment information. The split violin plot shows the distributions of reconstruction accuracy using models trained on data from all other patients (Across Subject, in black) and all other electrodes from the same patient (Within Subject, in gray). \textbf{B.  Distributions of correlation coefficients.} Since all patients participated in two experiments in dataset 2, we limit our analyses to task specific data.  The violin plots compare distributions of reconstruction accuracy using data from the same task (Within Experiment), the other task (Across Experiment), and from both tasks (All Experiment), using models trained on data from all other patients (Across Subject, in black) and all other electrodes from the same patient (Within Subject, in gray).}
\label{fig:supplemental_1}
\end{figure}

\begin{figure}[ptb]
\centering
\includegraphics[width=\textwidth]{figs/supplemental_2}
\caption{\small \textbf{Reconstruction quality for experiments 1 and 2.} \textbf{A. Distributions
      of correlation coefficients.} For experiment 1, the split violin plots compare distributions of reconstruction accuracy using data from just experiment 1 (Within Experiment), just experiment 2 (Across Experiment), and from both experiments (All Experiment), using models trained on data from all other patients (Across Subject, in black) and all other electrodes from the same patient (Within Subject, in gray). \textbf{B.  Distributions
      of correlation coefficients.} For experiment 2, the split violin plots compare distributions of reconstruction accuracy using data from just experiment 2 (Within Experiment), just experiment 1 (Across Experiment), and from both experiments (All Experiment), using models trained on data from all other patients (Across Subject, in black) and all other electrodes from the same patient (Within Subject, in gray). }
\label{fig:supplemental_2}
\end{figure}

\begin{figure}[ptb]
\centering
\includegraphics[width=\textwidth]{figs/supplemental_3}
\caption{\small \textbf{Reconstruction quality for each experiment.} \textbf{A. \& C.  Distributions
      of correlation coefficients.}  Across all electrodes from all
    patients in the labeled experiment from dataset 2, the panel displays the distribution of correlations between the observed and reconstructed LFP data using models trained on data from all other patients (Across, in black) and all other electrodes from the same patient (Within, in gray). 
    \textbf{B. \& D. Correlation maps.}  The glass brain maps display the
    average correlation between the observed LFP data and the across-subjects model reconstructed data by location, for each labeled experiment.}
\label{fig:supplemental_3}
\end{figure}

\begin{figure}[ptb]
\centering
\includegraphics[width=\textwidth]{figs/supplemental_4}
\caption{\small \textbf{Reconstruction accuracy predicted by data
    features for each dataset.} \textbf{A.-F. \& G.-L.  Correspondence
    between average
    reconstruction accuracy
    and data feature for datasets 1 and 2.} The scatterplots display the z-transformed
  average patient correlation coefficient versus different features of
  the data (\textbf{A. & G.} total electrodes, \textbf{B. & H.}  proportion thresholded, \textbf{C. & I.} total recording time,
  \textbf{D. & J.}  number thresholded, \textbf{E. & K.}  number of sessions, and \textbf{F. & L.}  average session
  length). The one-dimensional histograms report the distribution of
  the z-transformed correlation coeffiencients and the associated data
feature.}
\label{fig:supplemental_4}
\end{figure}

\begin{figure}[ptb]
\centering
\includegraphics[width=\textwidth]{figs/supplemental_5}
\caption{\small \textbf{ Top 10\% most informative voxels.}
 Voxels in yellow indicate the intersections between
    the top 10\% most informative voxels in each brain map from Figure
    5, indicated by the white contours.}
\label{fig:supplemental_5}
\end{figure}
